---
title: "SQLite database for CAZy.org"
output:
    github_document:
        toc: true
---

```{r libraries, echo=F}
library(knitr)
library(rmarkdown)
opts_chunk$set(echo=T, warning=F, message=F)
```

This repository provides the information stored at http://www.cazy.org as
computer-readable SQLite database. For this, it reads information stored as HTML
tables on the website into tabular data interpretable by Python and makes the
information available as a SQLite database.

## Preparation of database file prior to usage

The repository comes with the pre-generated SQLite database file. Due to the
size limitations of GitHub, the SQLite database is compressed using
[bzip2](https://www.sourceware.org/bzip2/). In order to use it, we have to
decompress it locally using

```{bash, eval=F}
bunzip2 -k CAZyme_genomes.db.bz2
```

The decompressed SQLite database can be used like any regular SQLite database. 

## Structure of the SQLite database

Here, I will use the scripting language R to illustrate the content of the
SQLite database. For this, we will use the two R packages
[DBI](https://cran.r-project.org/web/packages/DBI/index.html) and
[RSQLite](https://cran.r-project.org/web/packages/RSQLite/index.html).

```{r}
# Check if packages for connecting to the database are available
if (!all(c("RSQLite", "DBI") %in% installed.packages())) {
    install.packages(c("DBI", "RSQLite"))
}
library(RSQLite)
library(DBI)
library(magrittr)
```

We can open a connection to the database using the function `dbConnect`:

```{r}
conn <- dbConnect(RSQLite::SQLite(), "CAZyme_genomes.db")
```

By using `dbListTables`, we get an overview of the tables present in the
database:

```{r}
dbListTables(conn)
```

### Table "genomes"

The table **"genomes"** contains the information which CAZyme genes are present
in each genome. It has four different columns:

  - *name*: the name of the genome
  - *protein*: the name of the protein
  - *family*: the CAZy enzyme classification
  - *refAcc*: the NCBI GenBank reference ID of the protein

An example of the output looks like follows:

```{r, echo = F}
dbReadTable(conn, "genomes") %>%
head(10)
```

### Table "taxids"

The table **"taxids"** contains the information on taxonomic information of the
genomes. It has three different columns:

  - *name*: the name of the genome
  - *taxid*: the NCBI taxonomy ID
  - *superkingdom*: the superkingdom category it was classified in on cazy.org

An example of the output looks like follows:

```{r, echo = F}
dbReadTable(conn, "taxids") %>%
head(10)
```
